%=================================================================%

Question 9
Return to the previous question. Brain volume for adult women is about 1,100 cc for women with a standard deviation of 75 cc. Consider the sample mean of 100 random adult women from this population. Around what is the 95th percentile of the distribution of that sample mean?
1112 cc
1223 cc
1130 cc
1110 cc
%=================================================================%

Question 10
Recall from an earlier question. Suppose that diastolic blood pressures (DBPs) for men aged 35-44 are normally distributed with a mean of 80 (mm Hg) and a standard deviation of 10. What's the probability that in a random sample of 5 subjects, 4 or more have DBPs more than 90?
50%
16%
0.29 %
1%
5%

%=================================================================%

Question 11
Consider two sample means, X¯1 and X¯2 from a the same normal population having mean zero and standard deviation σ. The first sample mean is based on n1 observations while the second is based on n2 observations. Take any percentile, say 100×α, for the distribution of sample means based on the two sample sizes excluding the median. What is the ratio of the two percentiles where it is the percentile of the first group divided by the second?
σn1−−√/n2−−√
1
n1−−√/n2−−√
n1/n2
n2/n1
n2−−√/n1−−√
σn2−−√/n1−−√
%=================================================================%

Question 12
You flip a fair coin 5 times, what's the probability of getting 4 or 5 heads?
About 19%
About 11%
About 6%
About 25%

%=================================================================%

Question 13
The respiratory disturbance index (RDI), a measure of sleep disturbance, for a specific population has a mean of 15 (sleep events per hour) and a standard deviation of 10. They are not normally distributed. Give your best estimate of the probability that a sample mean RDI of 100 people is between 14 and 16 events per hour?
Around 99%
Around 90%
Around 68%
Around 95%

%=================================================================%

Question 14
Consider a standard uniform density. The mean for this density is .5 and the variance is 1 / 12. You sample 1,000 observations from this distribution and take the sample mean, what value would you expect it to be near?
1 / 12
1/12
1/12−−√
1/12×1,000−−−−−−−−−√
0.5
%=================================================================%

Question 15
Consider a standard uniform density. The mean for this density is .5 and the variance is 1 / 12. You sample 1,000 sample means where each sample mean is comprised of 100 observations. You take the standard deviation of the 1,000 sample means. About what number would you expect it to be?
1/(12×100)
1/12×100−−−−−−−√
1 / 12
1/12×1,000−−−−−−−−−√

%=================================================================%

Question 16
Consider a standard uniform density. The mean for this density is .5 and the variance is 1 / 12. You sample 1,000 sample variances where each sample variance is comprised of 100 observations. You take the average of the 1,000 sample variances. What number would you expect that to be near?
1/12−−√
1/12×100−−−−−−−√
1/(12×100)
1/12
1/(12×1,000)
1/12×1,000−−−−−−−−−√
