Fisher's Test
========================================================

*Agresti (1990, p. 61f; 2002, p. 91) Fisher's Tea Drinker*

A British woman claimed to be able to distinguish whether milk or
tea was added to the cup first.  To test, she was given 8 cups of
tea, in four of which milk was added first.  The null hypothesis
is that there is no association between the true order of pouring
and the woman's guess, the alternative that there is a positive
association (that the odds ratio is greater than 1).

```{r}
TeaTasting <-
matrix(c(3, 1, 1, 3),
       nrow = 2,
       dimnames = list(Guess = c("Milk", "Tea"),
                       Truth = c("Milk", "Tea")))
TeaTasting
fisher.test(TeaTasting, alternative = "greater")
## => p = 0.2429, association could not be established
```
### Criminal convictions of like-sex twins
*Fisher (1962, 1970)*
```{r}
Convictions <-
matrix(c(2, 10, 15, 3),
       nrow = 2,
       dimnames =
       list(c("Dizygotic", "Monozygotic"),
            c("Convicted", "Not convicted")))
Convictions
```
### Fisher's Tests
```{r}
fisher.test(Convictions, alternative = "less")
fisher.test(Convictions, conf.int = FALSE)
fisher.test(Convictions, conf.level = 0.95)$conf.int
fisher.test(Convictions, conf.level = 0.99)$conf.int
```
### Job Satisfaction
```{r}
## A r x c table  Agresti (2002, p. 57) Job Satisfaction
Job <- matrix(c(1,2,1,0, 3,3,6,1, 10,10,14,9, 6,7,12,11), 4, 4,
dimnames = list(income = c("< 15k", "15-25k", "25-40k", "> 40k"),
                satisfaction = c("VeryD", "LittleD", "ModerateS", "VeryS")))
Job
fisher.test(Job)
fisher.test(Job, simulate.p.value = TRUE, B = 1e5)
```
