Overview[edit]
 
PopHealth is an open source software tool that automates population health reporting quality measures for Meaningful Use. popHealth integrates with a healthcare provider's electronic health record (EHR) system to produce summary quality measures on the provider's patient population. popHealth demonstrates how a healthcare provider can use popHealth to analyze the quality of care provided as part of their existing workflow. popHealth helps healthcare providers easily understand the logic behind clinical quality measures and easily identifies patients requiring follow-up care.
 
popHealth leverages the Continuity of Care Records from a provider’s EHR to produce the quality measure reports. Specifically, popHealth leverages the HITSP CCD/C32 XML and ASTM CCR XML standards as patient inputs.
 
For vendors, popHealth integrates with multiple EHR systems and scales as clinical quality measures are added. Vendors and healthcare providers using clinical repositories for source data may integrate with popHealth, or leverage the quality measure engine that drives the typical report calculation.
 
popHealth allows healthcare providers to better understand their patient population and gives them tools to improve the health of their patients by performing their own population care analysis. popHealth provides a streamlined mechanism for sending data on summary quality measures from individual providers to public health organizations. popHealth was also designed to operate within a provider’s infrastructure and use encryption to address patient privacy and security; supports pre-defined quality measure reports, and supports the ability to export the PQRI XML as a reporting artifact for the transmission of summary data.
 
Background[edit]
 
The popHealth project is supported by the Office of the National Coordinator (ONC), under the United States Department of Health and Human Services (HHS). The popHealth open source software is currently being designed and developed by the MITRE Corporation, using components of the MITRE free open source Laika EHR testing framework.
 
popHealth demonstrates how a provider can use the system to submit quality measures or public health data as part of their existing workflow. The transmission of summary quality data is simpler, less data intensive and more scalable, representing an alternative to traditional methods of data analysis and reporting.
 
popHealth empowers physicians to better understand their patient population and gives them tools to improve the health of their patients by performing their own population care analysis. The software also provides a streamlined mechanism for sending data on summary quality measures from individual providers to public health organizations.
 
It was part of $100K challenge grant at www.health2challenge.org/pophealth-tool-development-challenge/
