\documentclass[12pt]{article}

%opening
\usepackage{framed}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\begin{document}


\section{Statistics One : Week 9 Quiz}
\subsection*{Description}
This week, we turn to a classic study on helping behavior by Darley and Batson (1973). Simulated data are provided here. The study demonstrates that people’s likelihood of helping a person in distress depends largely on their level of haste—whether they were running early, on time, or late for an appointment—when they encounter him, rather than on whether they have been asked to reflect on a pro-helping message (the parable of the Good Samaritan) as opposed to a neutral message (occupational effectiveness). In this dataset, independent variables include Prime (1 = parable of the Good Samaritan; 2 = occupational effectiveness) and Haste (1 = early, 2 = on time, 3 = late). 

On their way to a nearby location, participants encounter a moaning individual in distress. The Helping variable provides a measure of how much they help, ranging from 0 to 6 with higher scores indicating greater helping.

\subsection*{Question 1}
What is the class of Haste and Prime in R?


\subsection*{Question 2}
After converting Haste and Prime to factors, run an ANOVA with both Haste and Prime as independent variables. Is the effect of Haste significant? (Yes/No)
No
Yes
\subsection*{Question 3}
Is the effect of Prime significant? (Yes/No)
\subsection*{Question 4}
Is the interaction significant? (Yes/No)

\subsection*{Question 5}
Save the ANOVA summary in a table and run Tukey's pairwise comparison on all group means. Do each level of Haste significantly differ from one another? (Yes/No)

\subsection*{Question 6}
What is the partial eta-squared value for the effect of Haste? (round to 2 decimal places).


\subsection*{Question 7}
What is the partial eta-squared value for the interaction? (round to 2 decimal places).


\subsection*{Question 8}
Let's now run simple effects of Prime at each level of Haste. At which level of Haste is the effect of Prime significant?
\begin{itemize}
\item[(i)] Late
\item[(ii)] On time
\item[(iii)] Early
\item[(iv)] All of the above
\end{itemize}
\subsection*{Question 9}
What is the partial eta-squared value for the effect of Prime when people were early? (round to 2 decimal places).


\subsection*{Question 10}
Which one of the following statements best illustrates the main finding of the study?

\begin{itemize}
\item[(i)] People are less likely to help others if they are early
\item[(ii)] All of the above
\item[(iii)] People are more likely to be primed to help others if they are early
\item[(iv)] People are more likely to help others after being primed to do so if they are early
\end{itemize}






\end{document}
