%---------------------------------------------------------------%
Question 1
Standard Operating Procedures (SOPs) are “detailed, written instructions to achieve uniformity of the performance of a specific function.” Which of the following statements is false regarding SOPs:
Your Answer		Score	Explanation
Operational metrics and explicit SOPs can help your study be transparent for external review and provide a framework for accountability within your study team.	Inorrect	0.00	This statement is true.
Authoring SOPs can help you define the milestones on your study timeline and reveal potential roadblocks.			
The development of SOPs while planning your study helps you define the operation and technical components of your project.			
Authoring SOPs is enough to ensure that your study is in regulatory or contractual compliance.			
Total		0.00 / 1.00	
%---------------------------------------------------------------%
Question 2
Which of the following guidelines for creating SOPs were presented in the lecture video. (Check all that apply)
Your Answer		Score	Explanation
Committing sufficient time to SOP development is a good idea and will help derive more value from your SOPs.	Inorrect	0.00	This statement is true.
SOPs define best practices and as such whether the the study personnel sign-off to them is irrelevant.	Correct	0.50	SOPs are not just documents. They reflect procedures that real people will need follow. Getting their feedback will help you refine them in way that makes them more likely to be followed.
SOPs are living and evolving documents. You may need to modify them as the need arises.	Correct	0.50	This statement is true.
SOP templates are not very helpful because all studies are entirely different.	Correct	0.50	You don’t need to re-invent the wheel every time. Many “good practice” operational steps apply to multiple projects.
Total		1.50 / 2.00	
%---------------------------------------------------------------%
Question 3
For a study that evaluates patient-reported compliance with medication, on which of the following electronic systems would you rely as sources of data?
Your Answer		Score	Explanation
Pharmacy system if some or all medications are dispensed by your institution’s pharmacy.	Inorrect	0.00	This systems documents whether and when the patients physically obtained their medications.
The lab system for monitoring blood drug level.	Correct	0.50	This study by design relies on patient reported medication compliance; however you may want to take the extra step to verify that claim by testing for medication blood levels.
Physician and nursing documentation systems from subsequent visits.	Inorrect	0.00	These systems will inform you whether the patients reported to their care providers that they in fact took their medications.
Order entry and/or prescription writing system (Rx Writer).	Inorrect	0.00	This system will be used to ascertain whether a prescription of medication was in fact ordered by a physician and when.
Total		0.50 / 2.00	
%---------------------------------------------------------------%
Question 4
Which of the following are study data-related problems that stem from using aggregation systems:
Your Answer		Score	Explanation
If a production system is heavily used in the clinic (such as a scheduling or a billing system), your query to it may affect or be affected by real-time usage volume.	Inorrect	0.00	When you query an aggregation systems, you are not directly querying the source systems which are affected by the hour-to-hour operations of the hospital.
The data is never integrated across multiple systems.	Correct	0.50	While data integration across all systems may not be complete, aggregation systems typically do integrate data from across multiple source systems.
You may not be getting complete data	Inorrect	0.00	The data aggregation processes may not necessarily acquire all of the data present in the production system. The medical record or research data warehouse for example may not contain many of the intermediate processing timestamps that are part of pharmacy prescription or lab sample management for example.
You may not be getting real time data	Inorrect	0.00	The aggregation processes that transfer data from source to aggregation systems may not be running in real time (sometimes once every day or even less frequently).
Total		0.50 / 2.00	
%---------------------------------------------------------------%
Question 5
Electronic Data Capture systems can log all data viewing and manipulation events to create an audit trail. Why are audit trails important? (Check all that apply)
Your Answer		Score	Explanation
Audit trails are really only useful in limited clinical research scenarios like drug randomized controlled trials	Correct	0.25	
It is important to trace the history of every data point collected and changed throughout the lifetime of the study.	Inorrect	0.00	
Audit trails are no longer required or even needed for research studies	Correct	0.25	
An audit trail is often necessary to fulfill regulatory compliance requirements set by external agencies.	Inorrect	0.00	
Total		0.50 / 1.00	
Question Explanation

Tracing the history of every data point collected and changed throughout the lifetime of the study allows you to assure that authenticity and security of data were preserved and can help you improve and monitor the quality of the data. Maintaining audit trails may (sometimes) be an explicit requirement by external funding agencies, but that alone is not the main reason they are important.
%---------------------------------------------------------------%
Question 6
Electronic Data Capture systems would allow study personnel to enter research data into the database via Case Report Forms (CRFs). Which of the following functions are NOT desired in a CRF?
Your Answer		Score	Explanation
Data validation (preventing malformed dates for date fields, numeric range checking, etc) that prevents the users from entering values in given fields unless they conform to the data dictionary.			
Providing different field types (radio buttons, check boxes, drop down, free text fields, etc) based on the variable types.	Inorrect	0.00	
Displaying a human readable label or question associated with each field on the CRF.			
Displaying the variable names in the database that correspond to the questions on the CRF.			
Total		0.00 / 1.00	
Question Explanation

Variable names are often selected to conform to technical database naming schema. Showing them adds little value and can even confuse the users who will interact with the system via the CRF.
%---------------------------------------------------------------%
Question 7
An Electronic Data Capture system will need to allow users to export the study data. Which of the following statements do you agree with based on the content of the lectures? (Check all that apply)
Your Answer		Score	Explanation
An EDC system should be able to email the raw data directly to anyone on the internet that the study owner chooses.	Correct	0.50	Email is a good way to notify collaborators that they can obtain the data, but it is not a secure method for sending the data itself.
Date-shifting is one way that you can de-identify data upon export.	Correct	0.50	
The EDC system can help preserve subject privacy by giving the end-users a means to de-identify data upon export.	Inorrect	0.00	
The data exported from the system will most likely need to be analyzed. Therefore the systems should have capacity to export the data in formats that can be consumed by most data analysis software.	Inorrect	0.00	
Total		1.00 / 2.00	
%---------------------------------------------------------------%
Question 8
Electronic Data Capture systems can empower study investigators to create and manage their own projects without relying on software programmers. Which of the following features reflect that?
Your Answer		Score	Explanation
Capacity for study team members to create the underlying database and CRF elements using a metadata file that they can re-use or share with collaborators.	Inorrect	0.00	
Capacity for study team members to set up projects, but only if they document the purpose of their data capture study using a lengthy and very detailed questionnaire.	Correct	0.50	This defeats the purpose of user empowerment and ease-of-use features of EDC systems.
Capacity for study team members to create the underlying database and CRF elements using an intuitive (point and click) interface.	Inorrect	0.00	
Project owner ability to add and remove specific users and assign them permissions that are commensurate with their role in the study	Correct	0.50	
Total		1.00 / 2.00
%---------------------------------------------------------------%
