Question 1
Which of the following are advantages of using surveys to collect data directly from research participants? (Check all that apply)
Your Answer		Score	Explanation
YES Participant responses can be completely anonymous.	Inorrect	0.00	
YES Surveys help you avoid interviewer bias especially around sensitive topics.	Correct	0.20	
YES Direct data entry (especially if the survey instrument is electronic) eliminates data entry errors by study personnel.	Correct	0.20	
In the United States, you do not need IRB approval to conduct anonymous survey studies.	Correct	0.20	Even if you qualify for IRB exemption, you will still need to go through IRB approval mechanism in the United States.
If survey instructions are not clear, it is not a big problem, because survey participants can ask for clarification.	Correct	0.20	
Total		0.80 / 1.00	
%---------------------------------------------------------%
Question 2
Which of the following recommendations about survey questions were made in the video lectures? (Check all that apply)
Your Answer		Score	Explanation
Sometimes you and or your colleagues can think of interesting things to ask your respondents that are outside the scope of your specific aims. Adding those questions to your survey will give you additional information without any negative impact.	Correct	0.25	Additional questions may cause respondent fatigue. Your participants may not finish the survey or may not put enough effort into answering all questions including those that are needed for your specific aims.
You can condense questions and minimize respondent fatigue by creating compound questions that ask about multiple concepts which can be combined with "AND" or "OR".	Correct	0.25	
YES It is advisable to use simple grammar and neutral non-biased terms.	Inorrect	0.00	
YES Avoid acronyms and technical jargon unless it is appropriate for your audience.	Inorrect	0.00	
Total		0.50 / 1.00	
%----------------------------------------------------------%
Question 3
Recall bias ... (check all that apply)
Your Answer		Score	Explanation
Is due to people's tendency to lie about events that happened a long time ago.	Correct	0.25	
Is a poorly understood phenomenon.	Correct	0.25	
YES Can be minimized when the recall questions address a shorter and more recent time period.	Inorrect	0.00	
YES Can be minimized for the recent past using short series of related questions.	Inorrect	0.00	
Total		0.50 / 1.00	
%----------------------------------------------------------%
Question 4
Which of the following statements about survey questions is true based on the video lectures. (Check all that apply.)

Your Answer		Score	Explanation

YES Branching logic is one way to increase the validity of your data.	Correct	0.25	True. Branching logic steers the participants away from entering invalid response combinations to different questions e.g. providing pregnancy history while entering the sex at birth as "male."
Having branching logic in your survey increases survey fatigue.	Correct	0.25	False. By hiding unnecessary questions from the participants, branching logic minimizes the number of questions that they have to ponder.
YES Visual Analog Scales (VAS) can be represented as slider questions when administered online. VAS minimize respondent perception and memory bias.	Inorrect	0.00	True. See Data Collection Using Surveys 3 video.
Likert Scale questions should be balanced, should have equal distance in the gaps between the choices, and should have an even number of choices.	Inorrect	0.00	Having an even number of choices is one way to force the respondents to pick a non-neutral answer, but it is not a requirement for Likert Scales.
Total		0.50 / 1.00	
%----------------------------------------------------------%
Question 5
When you assemble your data collection instrument which of the following actions are likely to annoy, overwhelm, or discourage your participants?

Your Answer		Score	Explanation

Giving detailed explanation of the steps you are taking to ensure their anonymity in anonymous surveys.			
NO Mentioning up-front all the sources of information they will need to access to be able to answer your survey.	Inorrect	0.00	
Include ample information at the beginning of the survey explaining to survey respondents the intent and purpose of the survey.			
Don't reveal the real amount of time that it will take to fill the survey. Always under-represent the real amount of time to trick your respondents into starting the survey.			
Total		0.00 / 1.00	
%----------------------------------------------------------%
Question 6
It is important to perform pilot testing of your surveys by a small group of participants outside your group who represent your target population. From the data perspective this pilot testing is important because... (check all that apply)

Your Answer		Score	Explanation

YES You may find some invalid or inconsistent data in your pilot responses and can anticipate and correct that in your instrument.	Inorrect	0.00	
YES You may find out that many of the respondents do not have information at hand to answer some of your questions and you will therefore be informed about how complete your survey responses may be.	Correct	0.25	
NO If you're lucky, you may end up using this sample population's data to draw conclusions about your study hypothesis thereby saving the cost of the larger study.	Inorrect	0.00	Pilot testing is not typically conducted with the same rigorous criteria and parameters that should be in place for your actual study. This would make the resulting data non-suitable for study conclusions.
YES The pilot testing may inform you whether or not there will be variance in question responses (e.g. in Likert Scales)	Inorrect	0.00	
Total		0.25 / 1.00	
%----------------------------------------------------------%
Question 7
The advantages of using a de-centralized data hosting model for multi-center studies are (check all that apply):
Your Answer		Score	Explanation
NO Unlike in the centralized data hosting model, sites cannot see data from other sites.	Inorrect	0.00	This is statement is generally not true because typically in the centralized data hosting model the data from each of the sites are sequestered from each other at the data coordinating center level.
YES The data collection sites retain the ability to store and manage the data locally.	Inorrect	0.00	
YES Data can be collected using the sites' local data models. This eliminates the need by the study personnel to extract, transform, and load (ETL) the data into a common standard for analysis.	Correct	0.33	Without either (1) collecting the data using a common standard or (2) transforming the local data into a common standard, the data collected from multiple centers cannot be integrated and analyzed for the study.
Total		0.33 / 1.00	
%----------------------------------------------------------%
Question 8
Which of the following considerations for form design require understanding of your local settings:
Your Answer		Score	Explanation
NO You should always use paper forms and not computers or tablets to avoid the appearance of wealth.	Correct	0.25	This may be true in some cases, but in other cases and when feasible, electronic data capture or data capture via cellular phones may be the better option.
YES The address or place of residence of transitory populations may require that you structure your data collection to capture that information more frequently.	Inorrect	0.00	
YES High illiteracy means that many participants cannot spell their name which could lead to different phonetic spellings by different study personnel	Inorrect	0.00	
YES Incomplete record keeping around birth may mean that participants cannot accurately or consistently report their exact day of birth	Correct	0.25	
Total		0.50 / 1.00
%----------------------------------------------------------%
Question 9
Which of the following statements are true about Picture Archive and Communication System (PACS) (check all that apply)
Your Answer		Score	Explanation
NO Due to it's large memory capacity a PACS is the ideal machine to run complex image analysis algorithms.	Correct	0.25	PACSs are optimal for archival and transfer not for analysis. Images need to be downloaded from PACS into other machines were analysis software is run locally.
YES PACSs typically store images using the DICOM (Digital Imaging and Communications in Medicine) file format.	Inorrect	0.00	
NO PACSs are used for transferring image files between scanners where these images are eventually archived.	Inorrect	0.00	Scanners contain relatively little storage space. The final archival of imaging is typically in a PACS not in scanners.
YES Provides a read-only repository for storage and access of acquired imaging data.	Inorrect	0.00	
Total		0.25 / 1.00	
%----------------------------------------------------------%
Question 10
Which of the following hypothetical terminologies that would be used for an emergency department's registry of fractures has the highest granularity?
Your Answer		Score	Explanation
A terminology composed of two parts: the first term would be picked from a list of all the anatomical names of human bones and the second term would be picked from a controlled list of terms that describe the types of bone fractures: simple fracture, compound fracture, etc.			
NO A list of a lay terms describing the body part where a fracture occured (neck, head, left arm, right arm, left thigh, etc.)	Inorrect	0.00	
A list of all the anatomical names of human bones to indicate the specific bone where a fracture occurred.			
Total		0.00 / 1.00	




